---
title: "OWCProjectHockey"
author: "Song Qian"
date: "4/1/2019"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("FrontMatter.R")
load("owc_data.RData")
```
## Introduction

In this document, we present R code for fitting the Bayesian hockey stick model and the hierarchical hockey stick model.  We use the smoothed hockey stick model to improve numerical stability.  The smoothed hockey stick model is a nonlinear regression problem. Although a simple nonlinear least squares method would be sufficient, we include the Bayesian nonlinear code here as the starting point for the Bayesian hierarchical model.

## Fitting the Smoothed Hockey Stick Model

We fit the hockey stick model code without considering multilevel structure, a Bayesian nonlinear regression problem.

Using the five-day moving average of TP loading to OWC as the predictor variable.

First, all data (without considering seasons).  The following code chunk includes a function for generating Stan model input data and initial values, and the Stan model.  For all our Bayesian models, we assume that the slope of the first line segment is 0.

```{r}

## input function -- organizing input data to Bayesian model
stan.in0 <- function(infile=owc_daily, x="maTPLD05", y="TP_ef",
                    n.chains=4, stdz=F){
    keep <- (infile[,x] > 0) & (infile[,y] >0)
    infile <- infile[keep & !is.na(keep),]
    x <- log(infile[,x])
    if (stdz){
      xmu <- mean(x)
      xsd <- sd(x)
      x <- (x-mean(x))/sd(x)
    } else {
      xmu <- 0
      xsd <- 1
      }
    y <- log(infile[,y])
    n <- dim(infile)[1]
    inits <- list()
    if (stdz) theta <- 0.04
    else theta <- 0.01*diff(range(x))
    bugs.data <- list(N=n, y=y, x=x, theta=theta, beta1=0)
    for (i in 1:n.chains)
    	inits[[i]] <- list(beta0=rnorm(1), delta=runif(1),
		                 	   phi=runif(1, range(x)[1], range(x)[2]),
                         sigma=runif(1))
    para <- c("beta0", "delta", "phi", "sigma")
    return(list(para=para, data=bugs.data, inits=inits,
                n.chains=n.chains, mux=xmu, sdx=xsd))
}

## the stan model
    stan_model <- "
	  data{
	  int N; //the number of observations
	  vector[N] y; //the response
	  vector[N] x;

	  real theta;
          real beta1;
	}
	parameters {
	  real beta0; //the regression parameters
	  /*real beta1;*/
	  real<lower=0> delta;
	  real phi; //change point
	  real<lower=0> sigma;
	}
	transformed parameters {
	  vector[N] mu;
	  for (i in 1:N)
	    mu[i] = beta0 + beta1 * (x[i]-phi) +
		    delta * theta *
			    log1p(exp((x[i]-phi)/theta));
	}
	model {
	  phi ~ cauchy(0,1);
	  beta0 ~ normal(0,5);
	  delta ~ normal(0,5);

	  sigma ~ cauchy(0, 1);

	  y ~ normal(mu, sigma);
   }
    "

```

All Stan models will be fit in the following steps:

- Creating input data file, using standardized predictor (TP loading).  The input data file includes the mean and standard deviation of TP loading data for later use.
- Compiling the Stan model to create a model object
- Updating (MCMC runs)
- Extracting Stan output
- Reconstructing (original, un-standardizing) relevant model parameters ($\phi$ and $\delta$)
- Saving the Stan fit object and extracted coefficients
- Summarizing and plotting

```{r}
## step 1 -- input data file
input.to.bugs <- stan.in0(n.chains=nchains)

## step 2 -- compile stan model
fit_compile <- stan_model(model_code = stan_model)

## step 3 -- MCMC run
fit <- sampling(fit_compile, data = input.to.bugs$data,
                init = input.to.bugs$inits,
                pars = input.to.bugs$para,
                iter = niters, thin=nthin,
                chains = input.to.bugs$n.chains,
                control = list(adapt_delta = 0.98, max_treedepth=25))
## load("owcTPma05.RData")
print(fit)

## step 4 -- extract and re-construct (if x is centered)
fit2coef <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permuted=T))))
fit2coef$delta <- fit2coef$delta/input.to.bugs$sdx
fit2coef$phi <- input.to.bugs$mux + input.to.bugs$sdx*fit2coef$phi
fit2coef_sum <- summary(fit2coef)

## step 5 -- saving
save(fit, fit2coef, file="owcTPma05.RData")

## step 6 -- Plotting (for poster)
  xx <- log(owc_daily$maTPLD05)
  yy <- log(owc_daily$TP_ef)

## Using ggplot2
## 1 generic hockey-stick model
packages(latex2exp)
  
  png(file="hockey_gg.png", width=480, height=300)
  ggplot(data.frame(x=seq(0, 20, 0.1), y=hockey(seq(0,20,0.1), 5, 2, 10)), aes(x=x, y=y))+
    geom_line() +
    annotate("text", x=4.5, y=20, label=TeX("$y=\\beta_0+\\beta_1 x, $"), adj=0)+
    annotate("text", x=7.5, y=20, label=TeX("$x \\leq \\phi$"), adj=0)+
    annotate("text", x=4.5, y=18, label=TeX("$y=\\alpha_0+\\alpha_1 x, $"), adj=0)+
    annotate("text", x=7.5, y=18, label=TeX("$x > \\phi$"), adj=0)+
    annotate("text", x=5, y=16, label="Or")+
    annotate("text", x=4.5, y=14, label=TeX("$y=\\beta_0+\\alpha_1 (x-\\phi) I(x-\\phi)$"), adj=0)+
    annotate("text", x=14, y=15, label=TeX("$\\alpha_1$"), adj=0)+
    annotate("text", x=5, y=5.5, label=TeX("$\\beta_1=0$"), adj=1)+
    scale_x_continuous(name="TP Loading", breaks=10, labels = TeX("$\\phi$")) +
    scale_y_continuous(name="TP Concentration", limits = c(0, 25), breaks=5, labels = TeX("$\\beta_0$"))
  dev.off()

## NADB plot
## nadb_path <- "C:/users/songq/EESwithR/data"
nadb_path <- "~/OneDrive/EESwithR/2ndEd/R/Data"

nadb <- read.csv(paste(nadb_path, "nadb.csv", sep="/"))

png(file="nadb_gg.png", width=480, height=300)
ggplot(nadb, aes(x=PLI, y=TPOut)) + geom_point(color="blue")+
  annotate(geom="text", x=0.025, y=7, label="From North American Treatment Wetland Database\n Knight et al (1992)", size=3, hjust=0 )+
  ylab(TeX("TP Effluent Concentration ($\\mu$g/L)"))+
  scale_x_continuous(name=TeX("TP Loading (g m$^{-2}$ yr$^{-1}$)"),
                     breaks=c(0.1, 1, 25, 100, 1000),
                     label=c("0.1,","1", "25", "100", "1000"), trans = "log")
dev.off()

## OWC combined
  xrange<-range(xx[is.finite(xx)], na.rm = T)
  xeas <- seq(xrange[1], xrange[2], length.out = 100)
  yrv <- summary(hockey_smooth(xeas, fit2coef[1], 0, fit2coef[2], fit2coef[3]))
  ggdf <- data.frame(x=xeas, y=yrv[,1],
                     ylower=yrv[,4], yupper=yrv[,8])
  pp <- ggplot(data.frame(x=xx, y=yy), aes(x, y))+
    geom_point(color="blue") + geom_line(data=ggdf, aes(x, y), lwd=1.5)+
    geom_ribbon(data=ggdf, aes(x=x, ymin=ylower, ymax=yupper), alpha=0.3)+
    scale_x_continuous(name = "TP Loading (ton/yr)", breaks = log(c(0.01,0.1,1,10,100)),
                       labels = c(0.01,0.1,1,10,100))+
    scale_y_continuous(name = "TP Effluent Concentration (mg/L)", breaks=log(c(0.01,0.05,0.1,0.5,1)),
                       labels=c(0.01,0.05,0.1,0.5,1))
  
##  png(file="combinedFigGG.png", width=2*4*120, height=2*2.6*120)
  tikz(file="OWCcomb.tex", width=4, height=3, standAlone=F)
  print(pp+theme(aspect.ratio = 0.65))
  dev.off()
```


The model can also be fit with the predictor standardized so that data from multiple wetlands can be combined.

## Multilevel/Hierarchical Smoothed Hockey Stick Model

The four model parameters are allowed to vary by Season or another grouping variable.

```{r}
stan.in1 <- function(infile=owc_daily, x="maTPLD05", y="TP_ef",
                    grp="Season", n.chains=4, stdz=F){
    keep <- (infile[,x] > 0) & (infile[,y] >0)
    infile <- infile[keep & !is.na(keep),]
    x <- log(infile[,x])
    if (stdz) {
      xmu <- mean(x)
      xsd <- sd(x)
      x <- (x - xmu)/xsd
    } else {
      xmu <- 0
      xsd <- 1
    }
    y <- log(infile[,y])
    gr <- as.numeric(ordered(infile[,grp]))
    n <- dim(infile)[1]
    R <- max(gr)
    inits <- list()
    if (stdz) theta <- 0.04
    else theta <- 0.01*diff(range(x))
    bugs.data <- list(N=n, R=R, region=gr, y=y, x=x,
                      theta=theta, beta1=0)
    for (i in 1:n.chains)
	inits[[i]] <- list(beta0=rnorm(1), delta=runif(1),
			   phi=runif(1, range(x)[1], range(x)[2]),
                           re0=rep(0, R), reD=rep(0,R),
                           reP=rep(0,R),
                           sigma=runif(1),
                           sigma0=runif(1),sigmaD=runif(1), sigmaP=runif(1)
                           )
    para <- c("B0", "De", "Ph",
	      "delB0", "delD", "delPh",
	      "sigma", "sigma0","sigmaD","sigmaP"
              )
    return(list(para=para, data=bugs.data, inits=inits,n.chains=n.chains,
                mux=xmu, sdx=xsd))
}

## Stan model

    stan_model2 <- "
	  data{
	  int N; //the number of observations
	  vector[N] y; //the response
	  vector[N] x;
	  int R;
	  int region[N];

	  real theta;
          real beta1;
	}
	parameters {
	  real beta0; //the regression parameters
	  real<lower=0> delta;
	  real phi; //change point

	  vector[R] re0;
	  vector[R] reP;
	  vector[R] reD;
	  real<lower=0> sigma;
	  real<lower=0> sigma0;
	  real<lower=0> sigmaD;
	  real<lower=0> sigmaP;
	}
	transformed parameters {
	  vector[N] mu;
	  for (i in 1:N)
	    mu[i] = (beta0+re0[region[i]]) +
		    (beta1) * (x[i]-(phi+reP[region[i]])) +
		    (delta+reD[region[i]]) * theta *
			    log1p(exp((x[i]-(phi+reP[region[i]]))/theta));
	}
	model {
	  phi ~ cauchy(0,1);
	  beta0 ~ normal(0,5);
	  delta ~ normal(0,5);

	  sigma ~ cauchy(0, 1);
	  sigma0 ~ cauchy(0, 1);
	  sigmaD ~ cauchy(0, 1);
	  sigmaP ~ cauchy(0, 1);

	  re0 ~ normal(0, sigma0);
	  reD ~ normal(0, sigmaD);
	  reP ~ normal(0, sigmaP);

	  y ~ normal(mu, sigma);
	}
  generated quantities {
    real Ph;
    vector[R] delPh;
    real B0;
    vector[R] delB0;
    real De;
    vector[R] delD;
    Ph = phi + mean(reP[]);
    B0 = beta0 + mean(re0[]);
    De = delta + mean(reD[]);
    for (i in 1:R){
      delPh[i] = reP[i] - mean(reP[]);
      delB0[i] = re0[i] - mean(re0[]);
      delD[i]  = reD[i] - mean(reD[]);
    }
  }
    "
```

Now repeating the steps for the multilevel model
```{r}
input.to.bugs <- stan.in1(n.chains=nchains, stdz=T, x="TPloadX")

fit <- stan_model(model_code = stan_model2)
fit2 <- sampling(fit, data = input.to.bugs$data,
                 init=input.to.bugs$inits,
	               pars = input.to.bugs$para,
                 iter=niters, thin=nthin,
                 chains=input.to.bugs$n.chains,
                 control = list(adapt_delta = 0.99, max_treedepth=20))
##load("owcTPma05SeasonSTDZ.RData")
print(fit2)

## processing stan output
## Input loading is standardized for pooling data from multiple sites

muxSN <- input.to.bugs$mux
sdxSN <- input.to.bugs$sdx
fit2coefSN <- rvsims(as.matrix(as.data.frame(extract(fit2, permuted=T))))
fit2coefSN$De <- fit2coefSN$De/input.to.bugs$sdx
fit2coefSN$delD.1 <- fit2coefSN$delD.1/input.to.bugs$sdx
fit2coefSN$delD.2 <- fit2coefSN$delD.2/input.to.bugs$sdx
fit2coefSN$delD.3 <- fit2coefSN$delD.3/input.to.bugs$sdx
fit2coefSN$delD.4 <- fit2coefSN$delD.4/input.to.bugs$sdx
fit2coefSN$Ph <- fit2coefSN$Ph*input.to.bugs$sdx + input.to.bugs$mux
fit2coefSN$delPh.1 <- fit2coefSN$delPh.1*input.to.bugs$sdx
fit2coefSN$delPh.2 <- fit2coefSN$delPh.2*input.to.bugs$sdx
fit2coefSN$delPh.3 <- fit2coefSN$delPh.3*input.to.bugs$sdx
fit2coefSN$delPh.4 <- fit2coefSN$delPh.4*input.to.bugs$sdx
save(fit2, fit2coefSN, muxSN, sdxSN, file="owcTPmaXSeasonSTDZ.RData")

fit2coefSN_sum <- summary(fit2coefSN)

xx <- log(owc_daily$maTPLD05)
yy <- log(owc_daily$TP_ef)
flow <- owc_daily$maTPLD05

df <- data.frame(x=xx, y=yy, fl=flow)
df <- df[!is.na(df$fl) & df$fl > 0,]
df$percentile <- cut(df$fl, breaks=quantile(df$fl, prob=seq(0.0, 1, 0.1)), labels=paste(seq(10,100, 10), "%", sep=""))

## using ggplot2

xrng <- range(df$x)
df2 <- data.frame(x = seq(xrng[1], xrng[2],length=50),
                  y = hockey_smooth(seq(xrng[1], xrng[2],length=50),
                                    fit2coefSN_sum[1,2], 0,
             ##fit2coefSN_sum[2,7] + fit2coefSN_sum[9,7],
                                    fit2coefSN_sum[2,2],
                                    fit2coefSN_sum[3,2],
                                    input.to.bugs$data$theta))
    
q <- ggplot(data=df, aes(x, y, col=percentile)) #, color=col)
q <- q + geom_point()
q <- q + geom_path(data=df2, aes(x=x, y=y, col=NULL)) +
    scale_x_continuous(breaks=log(c(0.01,0.1,1, 10, 75, 150)),
                       labels = as.character(c(0.01,0.1,1, 10, 75, 150)))+
    scale_y_continuous(breaks=log(c(0.01,0.05, 0.1, 0.25, 0.5)),
                       labels = as.character(c(0.01,0.05, 0.1, 0.25, 0.5)))+
    labs(x="TP Loading (ton/yr)", y="TP concentration (mg/L)")


pdf(file="colorflow.pdf", width=5, height=3.5)
print(q)
dev.off()

tikz(file="colorflow.tex", width=5, height=3.5, standAlone = T)
print(q)
dev.off()

tikz(file="seasonalFig.tex", height=5, width=6.5, standAlone = T)
par(mfrow=c(2,2), oma=c(3, 3, 0.25,1), mgp=c(1.25,0.1,0), las=1, tck=0.01)
par(mar=c(0.25, 1, 0.25, 0.25))
plot(xx,yy,type="n", xlab="", ylab="", axes=F)
  for (i in 1:200){
    j <- sample(1:length(fit2coefSN$B0),1, replace=F)
    curve(hockey_smooth(x, fit2coefSN$B0[j]+fit2coefSN$delB0.1[j], 0,
##                           fit2coefSN$B1[j]+fit2coefSN$delB1.1[j],
                           fit2coefSN$De[j]+ fit2coefSN$delD.1[j],
                           fit2coefSN$Ph[j]+fit2coefSN$delPh.1[j],
                           input.to.bugs$data$theta), add=T, col="grey")
  }
  temp <- owc_daily$Season=="Spring"
  text(log(0.01), log(0.5), "Spring")
  points(log(owc_daily$maTPLD05[temp]), log(owc_daily$TP_ef[temp]))
  curve(hockey_smooth(x, fit2coefSN_sum[1,2] + fit2coefSN_sum[4,2], 0,
##                         fit2coefSN_sum[2,7] + fit2coefSN_sum[9,7],
                         fit2coefSN_sum[2,2] + fit2coefSN_sum[8,2],
                         fit2coefSN_sum[3,2] + fit2coefSN_sum[12,2],
                      input.to.bugs$data$theta), add=T, lwd=2)
  axis(2, at=log(c(0.05,0.1,0.25,0.5)),
     labels=as.character(c(0.05,0.1,0.25,0.5)))
  box()
  
##  plot(log(TP_ef)~log(maTPLD05), data=owc_daily, type="n")
  par(mar=c(0.25, 0.25, 0.25, 1))
  plot(xx,yy,type="n", xlab="", ylab="", axes=F)
  for (i in 1:200){
    j <- sample(1:length(fit2coefSN$B0),1, replace=F)
    curve(hockey_smooth(x, fit2coefSN$B0[j]+fit2coefSN$delB0.2[j], 0,
##                           fit2coefSN$B1[j]+fit2coefSN$delB1.2[j],
                           fit2coefSN$De[j]+fit2coefSN$delD.2[j],
                           fit2coefSN$Ph[j]+fit2coefSN$delPh.2[j],
                           input.to.bugs$data$theta), add=T, col="grey")
  }
  temp <- owc_daily$Season=="Summer"
  text(log(0.01), log(0.5), "Summer")
  points(log(owc_daily$maTPLD05[temp]), log(owc_daily$TP_ef[temp]))
  curve(hockey_smooth(x, fit2coefSN_sum[1,2] + fit2coefSN_sum[5,2], 0,
##                         fit2coefSN_sum[2,7] + fit2coefSN_sum[10,7],
                         fit2coefSN_sum[2,2] + fit2coefSN_sum[9,2],
                         fit2coefSN_sum[3,2] + fit2coefSN_sum[13,2],
                      input.to.bugs$data$theta), add=T, lwd=2)
 axis(4, at=log(c(0.05,0.1, 0.25, 0.5)),
       labels=as.character(c(0.01,0.1,0.25,0.5)))
 box()
 
##  plot(log(TP_ef)~log(maTPLD05), data=owc_daily, type="n")
 par(mar=c(0.25,1,0.25,0.25))
  plot(xx,yy,type="n", xlab="", ylab="", axes=F)
  for (i in 1:200){
    j <- sample(1:length(fit2coefSN$B0),1, replace=F)
    curve(hockey_smooth(x, fit2coefSN$B0[j]+fit2coefSN$delB0.3[j], 0,
##                           fit2coefSN$B1[j]+fit2coefSN$delB1.3[j],
                           fit2coefSN$De[j]+fit2coefSN$delD.3[j],
                           fit2coefSN$Ph[j]+fit2coefSN$delPh.3[j],
                           input.to.bugs$data$theta), add=T, col="grey")
  }
  temp <- owc_daily$Season=="Autumn"
  text(log(0.01), log(0.5), "Autumn")
  points(log(owc_daily$maTPLD05[temp]), log(owc_daily$TP_ef[temp]))
  curve(hockey_smooth(x, fit2coefSN_sum[1,2] + fit2coefSN_sum[6,2], 0,
##                         fit2coefSN_sum[2,7] + fit2coefSN_sum[11,7],
                      fit2coefSN_sum[2,2] + fit2coefSN_sum[10,2],
                         fit2coefSN_sum[3,2] + fit2coefSN_sum[14,2],
                      input.to.bugs$data$theta), add=T, lwd=2)
axis(2, at=log(c(0.05,0.1,0.25,0.5)),
     labels=as.character(c(0.05,0.1,0.25,0.5)))
axis(1, at=log(c(0.01,0.1,1,10,75,150)),
       labels=as.character(c(0.01,0.1,1,10,75,150)))
box()

  ##  plot(log(TP_ef)~log(maTPLD05), data=owc_daily, type="n")
par(mar=c(0.25,0.25,0.25,1))
    plot(xx,yy,type="n", xlab="", ylab="", axes=F)
  for (i in 1:200){
    j <- sample(1:length(fit2coefSN$B0),1, replace=F)
    curve(hockey_smooth(x, fit2coefSN$B0[j]+fit2coefSN$delB0.4[j], 0,
##                           fit2coefSN$B1[j]+fit2coefSN$delB1.4[j],
                           fit2coefSN$De[j]+fit2coefSN$delD.4[j],
                           fit2coefSN$Ph[j]+fit2coefSN$delPh.4[j],
                           input.to.bugs$data$theta), add=T, col="grey")
  }
  temp <- owc_daily$Season=="Winter"
  text(log(0.01), log(0.5), "Winter")
  points(log(owc_daily$maTPLD05[temp]), log(owc_daily$TP_ef[temp]))
  curve(hockey_smooth(x, fit2coefSN_sum[1,2] + fit2coefSN_sum[7,2], 0,
##                         fit2coefSN_sum[2,7] + fit2coefSN_sum[12,7],
                         fit2coefSN_sum[2,2] + fit2coefSN_sum[11,2],
                         fit2coefSN_sum[3,2] + fit2coefSN_sum[15,2],
                      input.to.bugs$data$theta), add=T, lwd=2)
axis(4, at=log(c(0.05,0.1,0.25,0.5)),
     labels=as.character(c(0.05,0.1,0.25,0.5)))
axis(1, at=log(c(0.01,0.1,1,10,75,150)),
       labels=as.character(c(0.01,0.1,1,10,75,150)))
box()
mtext("TP loading (ton/yr)", 1, 0.75,T)
mtext("TP concentration (mg/L)", 2, 0.75, T, las=0)
  dev.off()
  
  ## producing a tablle:
  
write.csv(summary(exp(fit2coefSN[3] + fit2coefSN[12:15])), file="thresholds.csv")

```

