---
title: "Old Woman Creek Project -- Bayesian Models "
author: "Song Qian"
date: "March 11, 2019"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Introduction 

This Rmd file includes Bayesian updating code for the Old Woman Creek project. Data used include daily monitoring data from Old Woman Creek (divided into four seasons), Crane Creek, Coldwater Creek, and Olentangy.  The code in the previous Rmd files generate necessary input data and prior distributions for model parameters. 

```{r setup, include=FALSE}
## functions and setup
to2 <- function(x){
    ## conver a single digit integer to 0x
    return(ifelse (x<10, paste("0",x, sep=""), as.character(x)))
}

hockey_smooth <- function(x, beta0, beta1=0, delta, phi, theta=NULL)
{
    if (is.null(theta)) theta=0.01*diff(range(x))
    return(beta0 + beta1 * (x-phi) +
           delta * theta * log1p(exp((x-phi)/theta)))
}

line.plots <- function(est, se, Ylabel, Xlab, yaxis=2, Xlim=NULL,
                       rscale=F, At=NULL, Lab=NULL, V=0){
    if (is.null(Xlim)) Xlim <- range(c(est+2*se, est-2*se))
    n <- length(est)
    if(n != length(se))stop("lengths not match")
    plot(1:n, 1:n, xlim=Xlim, ylim=c(0.75, n+0.25),
         type="n", axes=F, xlab=Xlab, ylab="")
    if (rscale)
        axis(1, at=At, labels=Lab)
    else
        axis(1)
    axis(yaxis, at=1:n, labels=Ylabel, las=1)
    segments(y0=1:n, y1=1:n, x0=est-2*se, x1=est+2*se)
    segments(y0=1:n, y1=1:n, x0=est-1*se, x1=est+1*se, lwd=2.5)
    points(est, 1:n)
    if (!is.null(V)) abline(v=V, col="gray")
    invisible()
}

packages<-function(x, repos="http://cran.r-project.org", ...){
    x<-as.character(match.call()[[2]])
    if (!require(x,character.only=TRUE)){
        install.packages(pkgs=x, repos=repos, ...)
        require(x,character.only=TRUE)
    }
}

packages(tidyverse)

packages(ggplot2)
packages(rv)
packages(arm)
packages(lattice)
packages(tikzDevice)
packages(car)
packages(formatR)
packages(rstan)

rstan_options(auto_write = TRUE)
options(mc.cores = min(c(parallel::detectCores(), 8)))

nchains <-  min(c(parallel::detectCores(), 8))
niters <- 100000
nkeep <- 2500
nthin <- ceiling((niters/2)*nchains/nkeep)

```
## Importing Data


```{r read}
base <- getwd()
dataDIR <- paste(base, "Data", sep="/")

load("owc_data.RData")

```

Prior function from previous Rmd file

```{r}
## using the hierarchcial mdoel for the prior

prior <- function(fit, b0="B0", de="De", ph="Ph",
                  s0="sigma0",sD="sigmaD",sP="sigmaP",
                  n0=20, setn0=F){
    fit2SNprior <- rvsims(as.matrix(as.data.frame(rstan::extract(fit, permute=T))))
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==b0])
    Eb0 <- tmp$mean
    Vb0 <- tmp$sd^2
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==de])
    EDe <- tmp$mean
    VDe <- tmp$sd^2
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==ph])
    EPh <- tmp$mean
    VPh <- tmp$sd^2
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==s0]^2)
    Esigma0 <- tmp$mean
    Vsigma0 <- tmp$sd^2
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==sD]^2)
    EsigmaD <- tmp$mean
    VsigmaD <- tmp$sd^2
    
    tmp <- summary(fit2SNprior[names(fit2SNprior)==sP]^2)
    EsigmaP <- tmp$mean
    VsigmaP <- tmp$sd^2

    if (setn0) {
        alpha0 <- n0+1
        alphaD <- n0+1
        alphaP <- n0+1
    } else {
        alpha0 <- 2+Esigma0/Vsigma0
        alphaD <- 2+Esigma0/VsigmaD
        alphaP <- 2+Esigma0/VsigmaP
    }
    beta0 <- Esigma0*(alpha0-1)
    betaD <- EsigmaD*(alphaD-1)
    betaP <- EsigmaP*(alphaP-1)
    lambda0 <- Esigma0/Vb0
    lambdaD <- EsigmaD/VDe
    lambdaP <- EsigmaP/VPh
    ## limiting alpha + beta to be less than 1000
    while (alpha0+beta0 > 1000){
      alpha0 <- alpha0/10
      beta0 <- beta0/10
    }
    while (alphaD+betaD > 1000){
      alphaD <- alphaD/10
      betaD <- betaD/10
    }
    while (alphaP+betaP > 1000){
      alphaP <- alphaP/10
      betaP <- betaP/10
    }
    return(list(m0=Eb0, mD=EDe, mP=EPh,
                lmbd0=lambda0, lmbdD=lambdaD, lmbdP=lambdaP,
                al0=alpha0, alP=alphaP, alD=alphaD,
                bt0=beta0, btP=betaP, btD=betaD))
}

```

### Revised hockey stick model

The Stan model now includes a hyper-parameter for each model coefficients:

```{r}
stan_model3 <- "
	  data{
	  int N; //the number of observations
	  vector[N] y; //the response
	  vector[N] x; 

	  real theta;
          real beta1;

          real m0;
          real mD;
          real mP;

          real lmbd0;
          real lmbdD;
          real lmbdP;

          real al0;
          real alP;
          real alD;

          real bt0;
          real btP;
          real btD;

	}
	parameters {
	  real beta0; //the regression parameters
	  real<lower=0> delta;
	  real phi; //change point

	  real<lower=0> sigma;

    real mu0;
    real muD;
    real muP;

	  real<lower=0> sigma0sq;
	  real<lower=0> sigmaDsq;
	  real<lower=0> sigmaPsq;
	}
	transformed parameters {
	  vector[N] mu;
	  real<lower=0> sigma0;
	  real<lower=0> sigmaD;
	  real<lower=0> sigmaP;
	  
    sigma0 = sqrt(sigma0sq);
    sigmaD = sqrt(sigmaDsq);
    sigmaP = sqrt(sigmaPsq);
	  for (i in 1:N)
	    mu[i] = beta0 + beta1 * (x[i]-phi) +
		    delta * theta *
			    log1p(exp((x[i]-phi)/theta));
	}
	model {  
	  sigma ~ cauchy(0, 1)  ;
	  sigma0sq ~ inv_gamma(al0, bt0);
	  sigmaDsq ~ inv_gamma(alD, btD);
	  sigmaPsq ~ inv_gamma(alP, btP);
       
    mu0 ~ normal(m0, sigma0/sqrt(lmbd0));
    muD ~ normal(mD, sigmaD/sqrt(lmbdD));
    muP ~ normal(mP, sigmaP/sqrt(lmbdP));
          
	  phi ~ normal(muP, sigmaP); 
	  beta0 ~ normal(mu0, sigma0); 
	  delta ~ normal(muD, sigmaD); 

	  y ~ normal(mu, sigma);
	}
"

stan.in2 <- function(infile=owc_daily, x="maTPLD05", y="TP_ef",
                     grp=owc_daily$Quarter=="Q1", n.chains=4,
                     stdz=T,info=F, prrs = NULL){
    if (info & is.null(prrs)) stop("Need informative priors")
    infile <- infile[grp,]
    keep <- (infile[,x] > 0) & (infile[,y] >0)
    infile <- infile[keep & !is.na(keep),]
    x <- log(infile[,x])
    if (stdz){
      xmu <- mean(x, na.rm=T)
      xsd <- sd(x, na.rm=T)
      x <- (x - xmu)/xsd
    } else {
      xmu <- 0
      xsd <- 1
    }
    y <- log(infile[,y])
    n <- dim(infile)[1]
    if (info){
        m0 = prrs$m0
        mD = prrs$mD
        mP = prrs$mP
        lmbd0=prrs$lmbd0
        lmbdD=prrs$lmbdD
        lmbdP=prrs$lmbdP
        al0=prrs$al0
        alP=prrs$alP
        alD=prrs$alD
        bt0=prrs$bt0
        btP=prrs$btP
        btD=prrs$btD
    }else{
        m0 = 0
        mD = 0
        mP = 0
        lmbd0=1
        lmbdD=1
        lmbdP=1
        al0=2
        alP=2
        alD=2
        bt0=2
        btP=2
        btD=2
    }

    s0 <- sqrt(bt0/(al0-1))
    sD <- sqrt(btD/(alD-1))
    sP <- sqrt(btP/(alP-1))
    
    inits <- list()
    if (stdz) theta <- 0.04
    else theta <- 0.01*diff(range(x))
    bugs.data <- list(N=n, y=y, x=x,
                      theta=theta, beta1=0,
                      m0 = m0,  mD = mD, mP = mP,
                      lmbd0=lmbd0, lmbdD=lmbdD, lmbdP=lmbdP,
                      al0=al0, alP=alP, alD=alD,
                      bt0=bt0, btP=btP, btD=btD )
    for (i in 1:n.chains)
	inits[[i]] <- list(beta0=rnorm(1, m0, s0),
	                   delta=rnorm(1,mD,sD),
			   phi=runif(1, range(x)[1], range(x)[2]),
			   sigma=runif(1), sigmaPsq=runif(1), sigmaDsq=runif(1),
                           sigma0sq=runif(1),
                           mu0=rnorm(1, m0,s0), 
			   muD=rnorm(1, mD, sD),
                             muP=rnorm(1, mP,sP))
    para <- c("beta0", "delta", "phi","sigma",
              "mu0","muD","muP", "sigma0","sigmaD","sigmaP"
              )
    return(list(para=para, data=bugs.data, inits=inits,n.chains=n.chains,
                mux=xmu, sdx=xsd))
}

load("owcTPma05SeasonSTDZ.RData")
fit2coefSN <- rvsims(as.matrix(as.data.frame(rstan::extract(fit2, permuted=T))))

prr <- prior(fit2)

```
# Bayesian Sample Size Estimation

Using Monte Carlo simulation, we draw random samples of various sizes based on existing estimates of the hockey stick model parameters. For each set of samples, another model is fit to estimate model parameters.  The results will be presented as a plot of the estimated threshold (with its 90% credible interval) as a function of sample size.  We use the Coldwater Creek data as an example.

```{r}
## input file:
stan.inSZ <- function(x, y, n.chains=4, prrs = prr, sz){
        n <- length(x)
        m0 = prrs$m0
        mD = prrs$mD
        mP = prrs$mP
        lmbd0=prrs$lmbd0
        lmbdD=prrs$lmbdD
        lmbdP=prrs$lmbdP
        al0=prrs$al0
        alP=prrs$alP
        alD=prrs$alD
        bt0=prrs$bt0
        btP=prrs$btP
        btD=prrs$btD
  
        s0 <- sqrt(bt0/(al0-1))
        sD <- sqrt(btD/(alD-1))
        sP <- sqrt(btP/(alP-1))
    
        inits <- list()
        theta <- 0.04
        bugs.data <- list(N=n, y=y, x=x,
                          theta=theta, beta1=0,
                          m0 = m0,  mD = mD, mP = mP,
                          lmbd0=lmbd0, lmbdD=lmbdD, lmbdP=lmbdP,
                          al0=al0, alP=alP, alD=alD,
                          bt0=bt0, btP=btP, btD=btD )
        for (i in 1:n.chains)
        	inits[[i]] <- list(beta0=rnorm(1, m0, s0),
	                           delta=runif(1),
		 	                       phi=runif(1, range(x)[1], range(x)[2]),
                    			   sigma=runif(1), sigmaPsq=runif(1),
                        		 sigmaDsq=runif(1), sigma0sq=runif(1),
                             mu0=rnorm(1, m0, s0), 
                    			   muD=rnorm(1, mD,sD),
                             muP=rnorm(1, mP, sP))
    para <- c("beta0", "delta", "phi","sigma",
              "mu0","muD","muP", "sigma0","sigmaD","sigmaP")
    return(list(para=para, data=bugs.data,
                inits=inits,n.chains=n.chains, sz=sz))
}
##cwc data
tmp <- cwc_month$flow!=0
load("ColdCreek_STDZ.RData")
fitcoefCWC_temp <- summary(rvsims(as.matrix(as.data.frame(extract(fitCWC, permute=T))))) ## using standardized only
fit <- stan_model(model_code = stan_model3)

## sample size

sim_sz <- function(n=12, k=100, stanmodel=fit, prior_coef=fitcoefCWC_temp){
     fitcoefSZ_sim <- list(phi=rvsims(0), 
                           beta0=rvsims(0), 
                           delta=rvsims(0))
     for (i in 1:k){
        print(paste(i, "of", k, ", sample size =", n))
        x <- rnorm(n)
        mnY <- hockey_smooth(x, prior_coef[1,2] , 0,
                                prior_coef[2,2] ,
                                prior_coef[3,2] ,
                                0.04)
        y <- rnorm(n, mnY, prior_coef[4,2])

        input.to.stan <- stan.inSZ(x, y, sz=n)
        fit_sim <- sampling(fit, data = input.to.stan$data,   
                            init = input.to.stan$inits,
                            pars = input.to.stan$para,
                            iter = niters, thin=nthin, 
                            chains=input.to.stan$n.chains,
                            control=list(adapt_delta = 0.99, 
                                         max_treedepth=25))
        temp_sim <-  rvsims(as.matrix(as.data.frame(extract(fit_sim, permute=T))))
        fitcoefSZ_sim$beta0 <- fitcoefSZ_sim$beta0 + temp_sim[1]/k
        fitcoefSZ_sim$delta <- fitcoefSZ_sim$delta + temp_sim[2]/k
        fitcoefSZ_sim$phi <- fitcoefSZ_sim$phi + temp_sim[3]/k
     }
     return(fitcoefSZ_sim)
}

n <- (1:10)*12

smplsz_sim <- list()
for (j in 1:4)
  smplsz_sim[[j]] <- sim_sz(n=n[j])

for (j in 5:6)
  smplsz_sim[[j]] <- sim_sz(n=n[j])

for (j in 7:8)
  smplsz_sim[[j]] <- sim_sz(n=n[j])

for (j in 9:10)
  smplsz_sim[[j]] <- sim_sz(n=n[j])

save(smplsz_sim, file="CWC_SZ_sim.RData")
load("CWC_SZ_sim.RData")

```

Results are presented graphically by comparing the estimated model coefficients, especially the estimated threshold.

```{r}
j=0
sim_plot_phi <- NULL
for (i in n) {
  j=j+1
  sim_plot_phi <- rbind(sim_plot_phi, c(i, unlist(summary(smplsz_sim[[j]]$phi))))
}
sim_plot_phi <- as.data.frame(sim_plot_phi)

j=0
sim_plot_beta0 <- NULL
for (i in n) {
  j=j+1
  sim_plot_beta0 <- rbind(sim_plot_beta0, c(i, unlist(summary(smplsz_sim[[j]]$beta0))))
}
sim_plot_beta0 <- as.data.frame(sim_plot_beta0)

j=0
sim_plot_delta <- NULL
for (i in n) {
  j=j+1
  sim_plot_delta <- rbind(sim_plot_delta, c(i, unlist(summary(smplsz_sim[[j]]$delta))))
}
sim_plot_delta <- as.data.frame(sim_plot_delta)

tikz(file="sz_sim.tex", width=6, height=4, standAlone=F)
par(mfrow=c(1,3), mar=c(4, 4, 1, 1), mgp=c(1.25,0.125, 0), 
    las=1, tck=0.01)
line.plots(sim_plot_phi$mean, sim_plot_phi$sd, 
           Xlab="Threshold (ton/yr)",
           Ylabel=sim_plot_phi[,1],
           rscale=T,
           At=log(seq(0.3, 0.8, 0.1)),
           Lab=as.character(seq(0.3,0.8,0.1)),
           V=NULL)
box()

line.plots(sim_plot_beta0$mean, sim_plot_beta0$sd, 
           Xlab="Intercept",
           Ylabel=sim_plot_beta0[,1],
           rscale=T,
           V=NULL)
box()

line.plots(sim_plot_delta$mean, sim_plot_delta$sd, 
           Xlab="Slope change",
           Ylabel=sim_plot_delta[,1],
           rscale=T,
           V=NULL)
box()

dev.off()

tikz(file="sz_simPhi.tex", width=3, height=4.75, standAlone=F)
par(mar=c(4, 4, 1, 1), mgp=c(1.25,0.125, 0), 
    las=1, tck=0.01)
line.plots(sim_plot_phi$mean, sim_plot_phi$sd, 
           Xlab="Threshold (ton/yr)",
           Ylabel=sim_plot[,1],
           rscale=T,
           At=log(seq(0.3, 0.8, 0.1)),
           Lab=as.character(seq(0.3,0.8,0.1)),
           V=NULL)
box()
dev.off()

tikz(file="sz_simBeta0.tex", width=3, height=4.75, standAlone=F)
par(mar=c(4, 4, 1, 1), mgp=c(1.25,0.125, 0), 
    las=1, tck=0.01)
line.plots(sim_plot_beta0$mean, sim_plot_beta0$sd, 
           Xlab="Intercept",
           Ylabel=sim_plot[,1],
           rscale=T,
           V=NULL)
box()
dev.off()

tikz(file="sz_simDelta.tex", width=3, height=4.75, standAlone=F)
par(mar=c(4, 4, 1, 1), mgp=c(1.25,0.125, 0), 
    las=1, tck=0.01)
line.plots(sim_plot_delta$mean, sim_plot_delta$sd, 
           Xlab="Slope Change",
           Ylabel=sim_plot[,1],
           rscale=T,
           V=NULL)
box()
dev.off()


```

