---
title: "Old Woman Creek Hydrologic Modeling"
author: "Song Qian"
date: "7/14/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages<-function(x, repos="http://cran.r-project.org", ...){
  x<-as.character(match.call()[[2]])
  if (!require(x,character.only=TRUE)){
    install.packages(pkgs=x, repos=repos, ...)
    require(x,character.only=TRUE)
  }
}

RHome <- getwd()
dataDIR <- paste(RHome,"Data", sep="/")
packages(arm)
packages(lattice)
packages(UsingR)
packages(ggplot2)
packages(dplyr)
packages(reshape2)
packages(tikzDevice)

lm.plots <- function(lm.obj){
    obj1<-xyplot(fitted(lm.obj)~(fitted(lm.obj)+resid(lm.obj)),
                 panel = function(x, y,...) {
                     panel.xyplot(x, y,...)
                     panel.abline(0,1, col="red",...)
                     panel.loess(x,y, span=1.0,col="green",...)
                     panel.grid()
                 },ylab="Fitted",xlab="Observed")
    ## checking whether the predicted is in greement with the observed

    obj2<-qqmath(~resid(lm.obj),
                 panel = function(x,...) {
                     panel.grid()
                     panel.qqmath(x,...)
                     panel.qqmathline(x,...)
                 }, ylab="Residuals", xlab="Standard Normal Quantile"
                 )
    ## checking whether residuals are normally distributed

    obj3<-xyplot(resid(lm.obj)~fitted(lm.obj), panel=function(x,y,...){
        panel.grid()
        panel.xyplot(x, y,...)
        panel.abline(0, 0)
        panel.loess(x, y, span=1, col=2,...)
    }, ylab="Residuals", xlab="Fitted")
    ## checking for patterns in residuals (independence)

    obj4<-xyplot(sqrt(abs(resid(lm.obj)))~fitted(lm.obj), panel=function(x,y,...){
        panel.grid()
        panel.xyplot(x, y,...)
        panel.loess(x, y, span=1, col=2,...)
    }, ylab="Sqrt. Abs. Residuals", xlab="Fitted")
    ## checking whether the residuals have a constant variance

    obj5<-rfs(lm.obj, aspect=1)
    ## visualizing R^2

    obj6 <- xyplot(cooks.distance(lm.obj) ~ fitted(lm.obj),
                   panel=function(x,y,...){
                       panel.xyplot(x,y,...)
                       panel.grid()},
                   ylab="Cook's Distance", xlab="Fitted")
    ## checking for influential data points

    print(obj1, position = c(0.0, 0.0, 0.5, 1/3), more = T)
    print(obj2, position = c(0.5, 0.0, 1.0, 1/3), more = T)
    print(obj3, position = c(0.0, 1/3, 0.5, 2/3), more = T)
    print(obj4, position = c(0.5, 1/3, 1.0, 2/3), more = T)
    print(obj5, position = c(0.0, 2/3, 0.5, 1.0), more = T)
    print(obj6, position = c(0.5, 2/3, 1.0, 1.0), more = F)
    invisible()
}

```
## Introduction
This document includes R code used for processing data related to the
Old Woman Creek (OWC) project. The main objective of the project is to
develop a predictive model of water depth for evaluating the potential
impact of projected future climate change on Old Woman Creek
hydrology.  The model will be used in conjunction with biological
monitoring program to project the climate impact on Old Woman Creek
wetland ecosystem.

This document records detailed data processing and model building
steps. Main sources of data are 
	1. USGS daily flow (discharge) data from gauging station 04199155
	2. OWC on-site meteorological monitoring data
	3. OWC-Lake Erie connection monitoring data
	4. OWC water quality monitoring data for water depth
	
## Data and Data processing

Reading and processing barrier beach opning status data
```{r}
MouthOC <- read.csv(paste(dataDIR, 'MOUTH.OC.1981.21518.csv', sep="/"))

#Organizing table colums from original data set. Column 1 = mouth open
#or closed (open = 1, closed = 0). Column 2 is reorganizing the date.
nrs <- dim(MouthOC)[1]
ncs <- dim(MouthOC)[2]
OpnCls <- unlist(MouthOC[,3:ncs], use.name = F)
mnth <- rep(MouthOC$Month, ncs - 2)
mnth <- as.numeric(ordered(mnth, levels = month.name))
dy <- rep(MouthOC$Day, ncs - 2)
year <- rep(names(MouthOC)[3:ncs], each = nrs)
year <- substring(year, 2,5)

MouthOpC <- data.frame(open = OpnCls,
                      date = as.Date(paste(mnth, dy, year, sep = "/"),
                                     format = "%m/%d/%Y"))

#Now we need to insert the depth and flow columns from the OL dataset.
#Read data into R

#Read in OL data to R
Olcsv <- read.csv(paste(dataDIR, "OWCOLWQ1.csv", sep="/"), skip = 2) 
Olcsv$date <- as.Date(as.character(Olcsv$DateTimeStamp),
                      format="%m/%d/%y %H:%M")

#Read in USGS flow data to R
USGSflow <- read.csv(paste(dataDIR, "USGSflowBR.txt", sep="/"),
                     skip = 29, sep='')
USGSflow$date <- as.Date(as.character(USGSflow$datetime), format="%Y-%m-%d")

##Berlin for flow, OL for depth

##Combine files together into 1 dataframe
##First change columns names to organaize the
##Merge by date
OWCFlowDepth <- merge(MouthOpC, Olcsv, by = "date", all = TRUE)
OWCFlowDepth <- merge(OWCFlowDepth, USGSflow, by ="date", all = TRUE)
colnames(OWCFlowDepth)[colnames(OWCFlowDepth)=="X109528_00060_00003"] <-
    "USGSFlow" 
OWCFlowDepth$USGSFlow <- as.character(OWCFlowDepth$USGSFlow) 
#Remove and change 'Ice' to NA
OWCFlowDepth$USGSFlow[OWCFlowDepth$USGSFlow == "Ice"] <- NA
OWCFlowDepth$USGSFlow <-as.numeric(OWCFlowDepth$USGSFlow) 


#Melt and Cast 
#Reshaping data to calculate daily averages 

#Melt to new dataframe
OWCMelt <- melt(OWCFlowDepth[,c("date", "USGSFlow", "open", "Depth")],
               id = "date")
head(OWCMelt)
 
#Cast to appear with daily averages
OWCcast <- dcast(OWCMelt, date ~ variable, mean)
head(OWCcast)

```

## Exploratory Data Analysis

Exploratory plots for potential predictors and transformation
```{r}
#Log Flow&Depth, open/ closed panels with all years
xyplot(log(Depth) ~ log(USGSFlow) | factor(open), data = OWCcast)

xyplot(Depth ~ USGSFlow | factor(open), data = OWCFlowDepth,
       main = "Flow and Depth measurements")

## there are negative depths and depth > 5
```

### Weather Data
```{r}
#Inserting Weather data. Starts 2001 till now
OWCmet <- read.csv(paste(dataDIR, "OWCOWMET1.csv", sep="/"), skip = 2)
OWCmet$date <- as.Date(as.character(OWCmet$DateTimeStamp),
                       format="%m/%d/%y %H:%M")

#Melt and Cast to get daily averages and use only the variables needed
#Melt to new dataframe
 
OWCMetmelt <- melt(OWCmet, id = "date",
                   measure = c("ATemp", "WSpd", "Wdir", "TotPAR", "TotPrcp"),
                   na.rm=TRUE)
head(OWCMetmelt)

#BP and RH are read in as character instead of numeric

#Cast to appear with daily averages
OWCMetcast <- dcast(OWCMetmelt, date ~ variable, mean, na.rm=T)
head(OWCMetcast)

#Combine weather data with OWCFlowDepth and create a new dataframe with all that
OWCHydroMet <- merge(OWCcast, OWCMetcast, by = "date", all = TRUE,
                     na.rm = TRUE)
#

#Fixing for Depth and adding Months, Years, and Seasons column
#Months Column
OWCHydroMet$Month <- format(OWCHydroMet$date, "%b")

#Years column
OWCHydroMet$Year <- as.numeric(format(OWCHydroMet$date, "%Y"))

#Removing the NA in year
OWCHydroMet <- OWCHydroMet[!is.na(OWCHydroMet$Year),]

#Seasons Column
OWCHydroMet$Season <- "Winter"

OWCHydroMet$Season[OWCHydroMet$Month=="Mar"|
            OWCHydroMet$Month=="Apr"| 
            OWCHydroMet$Month=="May"]="Spring"
OWCHydroMet$Season[OWCHydroMet$Month=="Jun"|
            OWCHydroMet$Month=="Jul"| 
            OWCHydroMet$Month=="Aug"]="Summer"
OWCHydroMet$Season[OWCHydroMet$Month=="Sep"|
            OWCHydroMet$Month=="Oct"| 
            OWCHydroMet$Month=="Nov"]="Fall"

#Fixing Depth
OWCHydroMet$Depth[OWCHydroMet$Year == 2008] <-
    OWCHydroMet$Depth[OWCHydroMet$Year == 2008] + .31
OWCHydroMet$Depth[OWCHydroMet$Year > 2008] <-
    OWCHydroMet$Depth[OWCHydroMet$Year > 2008] +.26

#Depth cannot be greater than 5 m (16.2 ft)
OWCHydroMet$Depth[OWCHydroMet$Depth > 5 | OWCHydroMet$Depth < 0] = NA

```

## Exploratory Plots
```{r}
pairs(log(Depth) ~USGSFlow + ATemp + WSpd + Wdir + TotPAR +TotPrcp,
      data = OWCHydroMet)

## 2007 to now
pairs(log(Depth) ~USGSFlow + ATemp + WSpd + Wdir + TotPAR +TotPrcp,
      data = OWCHydroMet[OWCHydroMet$Year > 2007,])


xyplot(log(TotPrcp) ~ log(USGSFlow) | factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

xyplot(TotPrcp ~ USGSFlow | factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

xyplot(log(TotPrcp) ~ log(USGSFlow) | factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

xyplot(log(USGSFlow)~ log(TotPrcp)  | factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])
```

### Moving Average Function
The following function calculates moving averages of a time series. 
```{r}
movingAverage <- function(x, n=1, centered=FALSE) {
    if (centered) {
        before <- floor  ((n-1)/2)
        after  <- ceiling((n-1)/2)
    } else {
        before <- n-1
        after  <- 0
    }
    
    ## Track the sum and count of number of non-NA items
    s     <- rep(0, length(x))
    count <- rep(0, length(x))
    
    # Add the centered data 
    new <- x
    # Add to count list wherever there isn't a 
    count <- count + !is.na(new)
    # Now replace NA_s with 0_s and add to total
    new[is.na(new)] <- 0
    s <- s + new
    
    # Add the data from before
    i <- 1
    while (i <= before) {
        # This is the vector with offset values to add
        new   <- c(rep(NA, i), x[1:(length(x)-i)])

        count <- count + !is.na(new)
        new[is.na(new)] <- 0
        s <- s + new
        
        i <- i+1
    }

    # Add the data from after
    i <- 1
    while (i <= after) {
        # This is the vector with offset values to add
        new   <- c(x[(i+1):length(x)], rep(NA, i))
       
        count <- count + !is.na(new)
        new[is.na(new)] <- 0
        s <- s + new
        
        i <- i+1
    }
    
    # return sum divided by count
    s/count
}
```

Using the moving average function, we now create Columns of moving
averages for Precip, Radiation, Flow, and Air Temp.
```{r}
#Precip moving average
OWCHydroMet$maPrcp10 <- movingAverage(OWCHydroMet$TotPrcp, 10)
OWCHydroMet$maPrcp20 <- movingAverage(OWCHydroMet$TotPrcp, 20)
OWCHydroMet$maPrcp30 <- movingAverage(OWCHydroMet$TotPrcp, 30)

#Radiation moving average
OWCHydroMet$maRad10 <- movingAverage(OWCHydroMet$TotPAR, 10)
OWCHydroMet$maRad20 <- movingAverage(OWCHydroMet$TotPAR, 20)
OWCHydroMet$maRad30 <- movingAverage(OWCHydroMet$TotPAR, 30)

#USGSFlow moving average 
OWCHydroMet$maFlow10 <- movingAverage(OWCHydroMet$USGSFlow, 10)
OWCHydroMet$maFlow20 <- movingAverage(OWCHydroMet$USGSFlow, 20)
OWCHydroMet$maFlow30 <- movingAverage(OWCHydroMet$USGSFlow, 30)

#Temp moving average
OWCHydroMet$maTemp10 <- movingAverage(OWCHydroMet$ATemp, 10)
OWCHydroMet$maTemp20 <- movingAverage(OWCHydroMet$ATemp, 20)
OWCHydroMet$maTemp30 <- movingAverage(OWCHydroMet$ATemp, 30)

#Now OWCHydroMet has all these new moving avergae columns

```

Exploring season effects:
```{r}
#Depth v log(Flow) by Seasons, CLOSED mouth
xyplot(Depth ~ log(USGSFlow) | Season, subset = open==0,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#############
xyplot(Depth ~ log(USGSFlow) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007 & OWCHydroMet$Depth > 0 ,])

#OPEN mouth
xyplot(Depth ~ log(USGSFlow) | Season, subset = open==1,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#Depth v Precip by Seasons CLOSED mouth
xyplot(Depth ~ TotPrcp | Season, subset = open ==0,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#OPEN mouth
xyplot(Depth ~ TotPrcp | Season, subset = open ==1,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#Depth v Radiation by Seasons CLOSED mouth
xyplot(Depth ~ TotPAR | Season, subset = open ==0,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#OPEN mouth
xyplot(Depth ~ TotPAR | Season, subset = open ==1,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#Depth v Air Temp by Seasons ClOSED mouth
xyplot(Depth ~ ATemp | Season, subset = open ==0,
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

#OPEN mouth
xyplot(Depth ~ ATemp | Season, subset = open ==1,
       data = OWCHydroMet)

#Precip Plots

plot(Depth~maPrcp10, data = OWCHydroMet)
plot(Depth~maPrcp20, data = OWCHydroMet)
plot(Depth~maPrcp30, data = OWCHydroMet) 

#Flow
plot(Depth ~ maFlow10, data = OWCHydroMet)
plot(Depth ~ maFlow20, data = OWCHydroMet)
plot(Depth ~ maFlow30, data = OWCHydroMet) 


#RAD
plot(Depth ~ maRad10, data = OWCHydroMet)
plot(Depth ~ maRad20, data = OWCHydroMet)
plot(Depth ~ maRad30, data = OWCHydroMet) 

#Air Temp
plot(Depth ~ maTemp10, data = OWCHydroMet)
plot(Depth ~ maTemp20, data = OWCHydroMet)
plot(Depth ~ maTemp30, data = OWCHydroMet) 


xyplot(Depth~log(maFlow10) | Season * factor(open),
       data = OWCHydroMet)

xyplot(Depth~log(maFlow30) | Season * factor(open),
       data = OWCHydroMet)

xyplot(log(Depth)~log(USGSFlow) | Season * factor(open), data = OWCHydroMet)


xyplot(Depth~log(maFlow10) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])
xyplot(Depth~log(maFlow20) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])
xyplot(Depth~log(maFlow30) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])


xyplot(Depth~log(USGSFlow) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])
xyplot(Depth~log(USGSFlow) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007,])

##PRCP

xyplot(Depth~log(maPrcp10) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maPrcp20) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maPrcp30) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])

##RAD
xyplot(Depth~log(maRad10) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maRad20) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maRad30) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])

##TEMP
xyplot(Depth~log(maTemp10) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maTemp20) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
xyplot(Depth~log(maTemp30) | Season * factor(open),
       data = OWCHydroMet[OWCHydroMet$Year > 2007, ])
```


## Exploring Model Forms

Linear regression models are fit and compared. 
```{r}
#All variables
OWClm <- lm(log(Depth) ~ maFlow10 + maRad10 + maPrcp10 + maTemp10 +
                maFlow20 +maRad20 +maPrcp20+ maTemp20 + maFlow30 +
                maTemp30 + maPrcp30 + maRad30 + USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm)

lm.plots(OWClm)

## removing least significant predictors
OWClm1 =lm(log(Depth) ~ maFlow10 + maRad10 + maTemp10 + maFlow20 +
               maRad20 + maTemp20 + maFlow30 + maTemp30 + maPrcp30 +
               maRad30 +USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm1)
lm.plots(OWClm1)

OWClm2 = lm(log(Depth) ~maFlow10 + maRad10 + maTemp10 + maFlow20 +
                maRad20 + maTemp20 + maTemp30 + maPrcp30 + maRad30 +
                USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm2)

OWClm3 = lm(log(Depth) ~maFlow10 + maRad10 + maFlow20 +maRad20 +
                maTemp20 + maTemp30 + maPrcp30 + maRad30 +USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm3)


OWClm4 =  lm(log(Depth) ~maFlow10 + maRad10 +maRad20 + maTemp20 +
                 maTemp30 + maPrcp30 + maRad30 +USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm4)

OWClm5 = lm(log(Depth) ~maFlow10 + maRad10 + maTemp20 + maTemp30 +
                maPrcp30 + maRad30 +USGSFlow,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm5)

OWClm6 = lm(log(Depth) ~maFlow10 + maRad10 + maTemp20 + maTemp30 +
                maPrcp30 + maRad30,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm6)

OWClm7 = lm(log(Depth) ~maFlow10 + maRad10 + maTemp20 +
                maTemp30 + log(maPrcp30) + maRad30,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm7)

OWClm8 = lm(log(Depth) ~log(maFlow10) + maRad10 + maTemp20 +
                maTemp30 + log(maPrcp30) + maRad30,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm8)

## add interaction
OWClm9 = lm(log(Depth) ~ log(maFlow10) + maRad10 + maTemp20 +
                maTemp30 + log(maPrcp30) + maRad30 +
                log(maPrcp30):log(maFlow10),
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm9)

##
OWClm10 = lm(log(Depth) ~ log(maFlow10) + maRad10 + maTemp30 +
                 log(maPrcp30) +log(maPrcp30):log(maFlow10) ,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm10)

## additive "open"
OWClm11 = lm(log(Depth) ~ log(maFlow10) + maRad10 + maTemp30 +
                 log(maPrcp30) +log(maPrcp30):log(maFlow10)+ open ,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm11)

## interactive "open"
OWClm12 = lm(log(Depth) ~ open *log(maFlow10) + open *maRad10 + open*maTemp30 +
                 open*log(maPrcp30) + log(maPrcp30):log(maFlow10) +
                 open:log(maPrcp30):log(maFlow10),
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm12)

## additive season
OWClm14 = lm(log(Depth) ~ Season + open *log(maFlow10) + open *maRad10 +
                 open*maTemp30 + open*log(maPrcp30) +
                 log(maPrcp30):log(maFlow10)+ open:log(maPrcp30):log(maFlow10),
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm14)

## removing insignificant interaction
OWClm12.1 = lm(log(Depth) ~ open * log(maFlow10) *log(maPrcp30) +
                   open * maTemp30 + open *maRad10,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm12.1)

OWClm14.1 = lm(log(Depth) ~ Season + open * log(maFlow10) *log(maPrcp30) +
                   open * maTemp30 + open *maRad10,
               data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                  OWCHydroMet$Depth > 0 &
                                  OWCHydroMet$Depth< 3.6 &
                                  OWCHydroMet$maFlow10 > 0,])

summary(OWClm14.1)

#Replace OPEN w/ SEASON
### MODEL 15 ###  (Model 12 replaced with season)
OWClm15 = lm(log(Depth) ~ Season * log(maFlow10) *log(maPrcp30) +
                 Season * maTemp30 + Season *maRad10,
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm15)

```

These models suggest that the final model should include both "open"
and "season" as factor variables.
```{r}
### Season and Open
OWClm16 = lm(log(Depth) ~ open * log(maFlow10) *log(maPrcp30) + 
               open * maTemp30 + 
               open * maRad10 + 
               Season * log(maFlow10) *log(maPrcp30) + 
               Season * maTemp30 + 
               Season * maRad10, 
             data = OWCHydroMet[OWCHydroMet$Year > 2007 &
                                OWCHydroMet$Depth > 0 &
                                OWCHydroMet$Depth< 3.6 &
                                OWCHydroMet$maFlow10 > 0,])
summary(OWClm16)
lm.plots(OWClm16)
```

The final model explains 33\% of the total variation in log depth. It
includes four 

## Predictive Model Example
```{r}

#Average Dataframe for prediction 
pred.data <- data.frame(open = rep(c(0,1), each= 4), 
                       Season = rep(c("Spring", "Summer", "Fall",
                                      "Winter"),2), 
                       maFlow10 = rep(mean(OWCHydroMet$maFlow10,
                                           na.rm = TRUE),8),
                       maPrcp30 = rep(mean(OWCHydroMet$maPrcp30,
                                           na.rm = TRUE),8), 
                       maTemp30 = rep(mean(OWCHydroMet$maTemp30,
                                           na.rm = TRUE),8), 
                       maRad10 = rep(mean(OWCHydroMet$maRad10,
                                          na.rm = TRUE), 8))

average.examp <- predict(OWClm16, new=pred.data, se.fit = TRUE,
                         interval = "prediction")$fit
cbind(pred.data, exp(average.examp))

```
The predictions suggest that the model is not accurate in predicting
winter depth

